\normalfont \normalsize \textsc{Northwestern University} \\ [25pt] \horrule{0.5pt} \\[0.4cm] \huge Masters Thesis \\ \horrule{2pt} \\[0.5cm]

### 2016-08-25 Gradschool/thesis/msthesis.tex
added memory concerns

@@

>>>>

\subsection{Memory Considerations}
Here we provide an overview of the memory resources required by the contributing structures involved in the likelihood calculation. This is a bookkeeping step necessary to select the correct sample array sizes given available hardware and tune the implementation to achieve optimal performance. 
The samples themselves are double precision arrays of length $N_s$. For $P$ params this equates to $8PN_s$ bytes of memory. The first computation in the routine is the generation of spherical harmonics, given a maximum value of $l$. Since there are $2l+1$ possible values of $m$ for each $l$, there are

\begin{align}
\sum_{l=0}^{L_{max}} (2l + 1) 
\end{align}

Total values of $m$ under a certain $L_{max}$. The partial sums are $S_L = (L_{max} + 1)^2$, so there are up to $16N_s(L_{max}+1)^2$ bytes of memory associated the spherical harmonics for a given value of $L_{max}$, requiring an extra factor of two to store complex numbers in double precision. Finally, we access to the complex conjugate of the spherical harmonics, this adds another factor of two bringing the total contribution to $32N_s(L_{max}+1)^2$ bytes.  
The antenna factor is also a complex number, but there is only a single value for each sample, adding only $16N_s$ bytes to the total memory usage, with an additional $16N_s$ for the conjugate. This is amounts to an additional $32N_s$ bytes. 
By far the greatest memory concern is the large time series block inside which the main calculation occurs. This block requires a row for each sample, multiplied by the number of modes for a total of $N_m N_s$ rows. Each of these rows holds $N_t$ elements that could, in the worst case, need to be padded by up to $T_{max} - 1$ zeros to enable proper time marginalization. Each element of this block is a double precision complex number. This equates to $16N_m N_s (N_t + T_{max} - 1)$ bytes of memory. An unfortunate shortcoming of PyCuda is that some simple operations can not be performed in place - one such operation is taking the real part of an array. This is due to the recasting of data types on the GPU itself. This means that for a brief portion of the calculation, an additional $50\%$ memory is needed raising the requirement to $24N_m N_s (N_t + T_{max} - 1)$ at peak for the main block.
The $U$ and $V$ crossterms are complex numbers that require $16N_s$ bytes each, for a total of $32N_s$. The second term of the likelihood, which is a combination of these, is a real number of the same length, adding $8N_s$ bytes to the total.
The sum of these contributions is the total number of bytes $B$ and is given by

\begin{align}
N_s(8P + 32(L_{max} + 1)^2 +  24N_m(N_t + T_{max} - 1) + 72)
\end{align}

Below we have tabulated $B$ for some values of $N_s$ to be used as reference when running the code.

@@

\begin{figure}[H!]
\hspace{-6cm}
\includegraphics[scale=0.5]{EMFollow.eps}
\end{figure}
